---
title: "PA1_template"
author: "Jesus"
date: "6/1/2021"
output:
  md_document: default
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# INIT

##### Loading and preprocessing the data. Show any code that is needed to:


1. Load the data (i.e. \color{red}{\verb|read.csv()|}read.csv())
2. Process/transform the data (if necessary) into a format suitable for your analysis
```{r  echo=TRUE}
library("data.table")
library(ggplot2)
library(tidyr)
library(lattice)

# Set working directory
setwd("C:/jesus/curso5/Practica1")
# If ref-folder doesnt exist, it must be created (a working directory must be selected first form R editor)
if (!file.exists("./ficheros")) {
        dir.create("./ficheros")
}

#Init variables
origen <-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
destino <- "./ficheros/ficheros.zip"
fileSource <- "./ficheros/activity.csv"

#Unzip file downloaded
if (!file.exists(destino )) {
        download.file(origen , destino , method = "curl")
        unzip(destino , overwrite = T, exdir = "./ficheros")
}
data_ini <- read.csv(fileSource,na.strings = "NA")
#Supress NA 's 
lines <- complete.cases(data_ini)
data_def <- data_ini[lines,]
head(data_def,4)
```


##### What is mean total number of steps taken per day?(For this part of the assignment, you can ignore the missing values in the dataset):

1. Calculate the total number of steps taken per day

```{r echo=TRUE}
#Sum steps by date
data_def_steps_day <- aggregate(steps ~ date, data=data_def, FUN=sum)
head(data_def_steps_day,14)

```

2. If you do not understand the difference between a histogram and a barplot, research the difference between them. Make a histogram of the total number of steps taken each day

```{r echo=TRUE}
#Histogram
hist(data_def_steps_day$steps, breaks=10, col="blue", main="Histogram total steps by day", xlab="Total   steps by day")
#dev.copy(png, file = "./plot1.png");
#dev.off();
```
3. Calculate and report the mean and median of the total number of steps taken per day

```{r echo=TRUE}
#Mean
mean(data_def_steps_day$steps)
#Median
median(data_def_steps_day$steps)
```


##### What is the average daily activity pattern?
1. Make a time series plot (i.e. \color{red}{\verb|type = "l"|}type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r echo=TRUE}
#aggregate steps by interval and plot the INFO
data_def_interval_5 <- aggregate(steps~ interval,data=data_def,FUN=mean)
plot(data_def_interval_5, type = "l", main="Average of steps by 5 minutes interval")
#dev.copy(png, file = "./plot2.png");
#dev.off();
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r echo=TRUE}
max_step_interval_5 <- data_def_interval_5[which.max(data_def_interval_5$steps),]
max_step_interval_5
```


##### Imputing missing values
Note that there are a number of days/intervals where there are missing values (coded as \color{red}{\verb|NA|}NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with \color{red}{\verb|NA|}NAs)
```{r echo=TRUE}
sum(is.na(data_ini$steps))
```


2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
-->Option:  or the mean for that 5-minute interval

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r echo=TRUE}
data_without_na <- data_ini

for (i in 1:nrow(data_without_na) ){
 #If steps are NA
   if ( is.na(data_without_na[i,1])){
     #We substitute the na value with the mean of the interval using previous dataset data_def_interval_5
   
    value <- data_def_interval_5[data_def_interval_5$interval==data_without_na[i,3],2]
    data_without_na[i,1] <- value
     
  }
}

```

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r echo=TRUE}
data_def_steps_day_wna <- aggregate(steps ~ date, data=data_without_na, FUN=sum)
hist(data_def_steps_day_wna$steps, breaks=10, col="blue", main="Histogram  total  steps by day without NULLS", xlab="Total   steps in a day")
#dev.copy(png, file = "./plot3.png");
#dev.off();
mean(data_def_steps_day_wna$steps)
median(data_def_steps_day_wna$steps)
```




##### Are there differences in activity patterns between weekdays and weekends?
For this part the \color{red}{\verb|weekdays()|}weekdays() function may be of some help here. Use the dataset with the filled-in missing values for this part.

1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r echo=TRUE}
#field date with date type content
data_without_na$date <- as.Date(data_without_na$date, "%Y-%m-%d")
#New field with the day of the week associated
data_without_na$day <- weekdays(data_without_na$date)
#New field to indicate if it is a weekday or weekend

data_without_na$daytype[(data_without_na$day == "Saturday" | data_without_na$day == "Sunday")] <- "weekend"
data_without_na$daytype[!(data_without_na$day == "Saturday" | data_without_na$day == "Sunday")] <- "weekday"


data_without_na$interval <- factor(data_without_na$interval, exclude = "")
data_without_na$daytype <- factor(data_without_na$daytype, exclude = "")

head(data_without_na,5)

```


2. Make a panel plot containing a time series plot (i.e. \color{red}{\verb|type = "l"|}type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.

```{r echo=TRUE}
data_without_na_week <- aggregate(steps ~ interval + daytype, data=data_without_na, FUN=mean)
head(data_without_na_week,4)
names(data_without_na_week) <- c("interval", "day_type", "steps")

xyplot(steps ~ interval| day_type, data = data_without_na_week, xlab = " interval", type = c("l", "l"), layout = c(1, 2), ylab = "average number of steps", main = "average number of steps  all days")




#dev.copy(png, file = "./plot4.png");
#dev.off();

```


